\chapter[Internet]{The Internet {\footnotesize "Post cold-war modern times"}}

The internet is a fundamental communication technology for today's modern society. The thing that started as the \Gls{arpanet}\cite{wiki:ARPANET} back in the late nineteen sixties has evolved to become the core of today's globalization on Earth.

Many things, people and creations has come to be part of the internet we know today.

\section[SP]{Service Providers}

They provide the basic connection to the \tsq{internet} we know.

\begin{itemize}
    \item \itemhead{Provider Classes}
    \begin{itemize}
        \item \textbf{Tier 1:} The granddaddies which does not pay for \Gls{ip} transit traffic from any other providers. They typically operate on the grand scale of the world and will often maintain peerings with most other tier 1 providers.
        \item \textbf{Tier 2:} Pays for \Gls{ip} transit traffic from the tier 1 providers and will maintain a select number of peerings between each other to reach certain prefixes by a shorter path. The granddaddies are paid to be able to reach the parts of the internet the tier 2 providers cannot reach through peerings with other tier 2 providers and \Glspl{cdn}.
        \item \textbf{Tier 3:} These networks can be large fx. large corporations, smaller local providers who do \textit{not} have downstream \Gls{ip} transit customers of their own and no peering relation ships.\footnote{Conditions comparable to \Gls{ospf} stub networks}
    \end{itemize}
    \item \itemhead{Provider Types}
    \begin{itemize}
        \item
    \end{itemize}
    \item \itemhead{Point to remember}
    \begin{enumerate}
        \item
    \end{enumerate}
\end{itemize}

\section[IXP]{Internet Exchange}

\section[MPLS]{Multiprotocol Label Switching}

\section[BGP]{Border Gateway Protocol}

\subsection[MP-BGP]{Multipoint Border Gateway Protocol}

\section[EVPN]{Ethernet Virtual Private Network}
