\chapter{Layer 3}

\section{Routed Network}

\subsection{Administrative Distance}

\begin{table}[]
    \centering
    \resizebox{\textwidth}{!}{%
        \begin{tabular}{|l|l|}
            \hline
            \textbf{Routing Protocol} & \textbf{Administrative distance} \\ \hline
            Directly connected interface & 0 \\ \hline
            Static route out an interface & 1 \\ \hline
            Static route to next-hop address & 1 \\ \hline
            DMNR - Dynamic Mobile Network Routing & 3 \\ \hline
            EIGRP summary route & 5 \\ \hline
            External BGP & 20 \\ \hline
            Internal EIGRP & 90 \\ \hline
            IGRP & 100 \\ \hline
            OSPF & 110 \\ \hline
            IS-IS & 115 \\ \hline
            Routing Information Protocol (RIP) & 120 \\ \hline
            Exterior Gateway Protocol (EGP) & 140 \\ \hline
            On Demand Routing (ODR) & 160 \\ \hline
            External EIGRP & 170 \\ \hline
            Internal BGP & 200 \\ \hline
            Next Hop Resolution Protocol (NHRP) & 250 \\ \hline
            Floating Static Route (ex. DHCP-learned) & 254 \\ \hline
            Unknown (Others) & 255 \\ \hline
        \end{tabular}%
    }
    \caption{Cisco default administrative distances}
    \label{cisco-default-administrative-distances}
\end{table}

Always remember the following points for Cisco devices:\cite{Administ9:online}
\begin{itemize}
    \item An administrative distance of 255 will cause the router to remove the route from the routing table and not use it.
    \item Since IOS 12.2, the administrative distance of a static route with an exit interface is 1. Prior to the release of 12.2 it was in fact 0.
    \item Only the interface itself has an administrative distance of 0, since a route cannot have a distance of less than 1.
    \item Directly connected routes have an administrative distance of 0.
\end{itemize}

\newpage

\section{OSPF}

\newpage

\section{IS-IS}

\newpage

\section{EIGRP}

\newpage

\section{RIP}

\newpage

\section{Static}

\newpage

\section{BGP}

\wikicommons{BGP_FSM}

The protocol of the internet used since 1994.\cite{wiki:Border_Gateway_Protocol}
Currently based upon \rfc{4271} with updates following in \rfc{6286} \rfc{6608}, \rfc{6793}, \rfc{7606}, \rfc{7607}, \rfc{7705}.

\subsection{Properties}

\begin{itemize}
    \item Uses tcp/179 as \gls{dst} port
    \item Sends keep-alive message every 1 minute
    \item Keep-alive message is 19 byte long
\end{itemize}

Be ware if sessions are terminated immediately upon trying to establish connection. Try debugging following points.

\begin{itemize}
    \item tcp/179 is not open,
    \item random port 1023> is not open,
    \item incorrect peer-ip,
    \item incorrect peer-as.
\end{itemize}

\subsection{Route exchange}

Exchanging routes between routers is a reliant and tolerant manner is \glspl{bgp} 1-advantage over \gls{ospf}/\gls{isis}/\gls{rip}/\gls{eigrp}.

The sheer tuning and control mechanisms \gls{bgp} can offer is simply astounding. Route-maps is the key and access-lists just one option.

\subsubsection[Route-maps]{Route-maps mechanism}

Route-maps is used to target a select set of routes and either modify/add/remove attributes attached to the select route-set.

\begin{itemize}
    \item Routes can be aggregated between \glspl{as}.
    \item Properties can be changed on the fly by matching
    \begin{enumerate}[label={\alph*)}]
        \item \Gls{bgp} communities,
        \item \Gls{ip} prefix,
        \item \Gls{bgp} as-path, 
    \end{enumerate}
\end{itemize}

An simple example of using route-maps is

\begin{cisco}
ip prefix-list 1 permit 172.16.0.0/16 
ip prefix-list 2 permit 192.168.1.0/24 
!
route-map RED permit 10 
 match ip address prefix-list 1 
 set ip next hop 10.1.1.1
 continue 20              ! Continues to apply rules normally only
                          ! applied to prefix-list 2. To apply to
                          ! prefix-list 1, too.
                          ! Any attributes set in '20' will
                          ! override any set during '10'.
route-map RED permit 20 
 match ip address prefix-list 2 
 set ip next hop 10.2.2.2 ! Last rule overrides previous rules from
                          ! previous '10' rule-set.
\end{cisco}

When rules from a rule-set is chained together as shown above. The last rule will override all previous set values regarding the attribute being applied. In this case \texttt{next-hop} from 'permit 10' is overridden in 'permit 20'.

\subsection[States]{BGP States}

The states is the way \gls{bgp} handles peer/neighbor connection establishing. The \underline{playbook} so to speak.

\begin{enumerate}
    \item Idle: \gls{bgp} while initializing refuses all incoming connections. Will initiate \gls{tcp} connection to peer.
    \item Connect: Waits for \gls{tcp} connection. If \gls{tcp} is established goes to state OpenSent. If \gls{tcp} is \textit{un}successful ConnectRetry timer is started and then goes to Active state.
    \item Active: When ConnectRetry counter reaches 0 goes to state Connect.
    \item OpenSent: Sends \gls{msg} to remote node. Waits for reply \gls{msg} before going to OpenConfirm.
    \item OpenConfirm: Nodes exchange keepalive \glspl{msg} and goes to Established state if successful.
    \item Established: Nodes can now exchange KeepAlive, Updates, and Notification \glspl{msg}.
\end{enumerate}

\subsection[iBGP]{Internal Border Gateway Protocol}

\gls{ibgp} is running \gls{bgp} within the same \gls{as} between routers. Much like running a general \gls{igrp} in the network.

Tradition one has to be fearful of creating \textit{routing loops} in the network. \glspl{bgp} mechanism for this is using either \begin{mylist} \item Full Mesh, or \item \glspl{rr} \end{mylist}.

Problems by running \textit{Full Mesh} is the formula of \[ iBGPsessions = n*(n-1)/2 \] \note{where $ n $ is the number \gls{ibgp} speakers} which results in scaling problems as \gls{ibgp} speakers are added to the \gls{as}.

\textit{\glspl{rr}} solves this problem by peering with all \gls{ibgp} speakers in the \gls{as}. All \gls{ibgp} speakers are then clients of the \glspl{rr}. This in turn helps maintainability by also advertising routes learnt from \gls{ibgp} clients to clients. Classic filtering/mathing route-maps/prefix-filters can be used to \textit{not} advertise all routes select group of clients from the \glspl{rr}.

\subsection[eBGP]{External Border Gateway Protocol}

\gls{ebgp} connections is inherently different from \gls{ibgp} connections. Some assumptions are made such as
\begin{enumerate}
    \item a \gls{ttl} of 1 is the default\footnote{Multi-hop \gls{ebgp} can thou be configured and therefore increase the max-\gls{ttl} value},
    \item distance is set to 20 compared to 200 for \gls{ibgp} routes,
    \item Next hop does \textit{not} change for \gls{ebgp} routes advertised to \gls{ibgp} neighbours \textit{by-default}\footnote{Often times it is necessary to tell a router to set itself as the next-hop before advertising to \gls{ibgp} neighbours}.
\end{enumerate}